---
title: "Chicago Data Explorer"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    social: menu
    author: 'Ryan Wang'
    theme: bootstrap
    vertical_layout: fill
    # logo: www/73F1665C-0FE6-4183-8AD1-E91DB8EFAFA4.png
runtime: shiny
---
<style>
:root{
  font-family: 'Open Sans';
}
.navbar {
  background-color:#c7e5f3;
  border-color:#c7e5f3;
}

.section.sidebar {
  top: 61px;
}

.navbar-brand {
  color:black!important;
  font-weight: 500;
  
.sidebar {
  font-weight: 400;
}
</style>   
```{r setup, include=FALSE}
library(leaflet)
library(spData)
library(dplyr)
library(ChicagoPackage)
library(tmap)
library(sf)
library(DT)
```

The Maps
=======================================================================
Column {data-width=350}
-----------------------------------------------------------------------
### Select Dataset and Attributes
```{r}
span(tags$i(h6("Data current as of June 2020.")), style="color:#045a8d")
selectizeInput("dataset", "Choose a dataset:", choices = c("Chicago Public Shools 2018", "Divvy", "CTA Stations", "Social Vulnerability Index"))
selectInput("variable", "Choose a variable:", choices = c("1", "2", "3"))
selectInput("by", "visualize by:", choices = c("Community Area", "2010 Cencus Tracts", "Police Beats", "Police Wards"))
selectInput("as", "show as:", choices = c("density dot", "dots", "Choropleth Map"))
sliderInput("year",
            "Year",
            min = 2001,
            max = 2018,
            step = 1,
            sep = "",
            value = 1)
```

### Download the Plot
```{r}
radioButtons("picformat", "Select the Image Format to Download", list("png", "jpeg"), "")

downloadButton(outputId = "down", label = "Download the Graph")
downloadHandler(filename = function() {
     paste('resultset-', Sys.Date(), '.csv', sep='')
   },
     content = function(file) {
     write.csv(subset_dataset(), file, row.names = FALSE)
   }
)
```

Output {data-width=650}
-----------------------------------------------------------------------
### Output Map
```{r}
# update region selections
  observeEvent(input$dataset, {
    if (input$dataset=="Divvy") {
      updateSelectInput(session = session, inputId = "variable", 
                        choices = c("Total Docks", "Docks in Service"))
      updateSelectInput(session = session, inputId = "by", 
                        choices = c("Community Area", "2010 Census Tracts", "Police Wards"))
      updateSelectInput(session = session, inputId = "as", 
                        choices = c("Choropleth Map", "Dots"))
    }
     if (input$dataset=="Chicago Public Shools 2018") {
      updateSelectInput(session = session, inputId = "variable", 
                        choices = c("School Type", "Primary Category", "Student Growth Rating", "Student Attainment Rating", "Culture Climate Rating", "Healthy School Certification", "School Survey Amibitious Instruction", "School Survey Safety", "Suspensions Per 100 Students Year 1 Percent"))
      updateSelectInput(session = session, inputId = "by", 
                        choices = c("NA"))
      updateSelectInput(session = session, inputId = "as", 
                        choices = c("Dots"))
    }
  }, ignoreInit = TRUE)

  # create dataframe with selected data
  reactive_db = reactive({
    if (input$dataset=="Divvy") { 
      db = divvy
    }
  })

renderLeaflet({
  leaflet(ca_boundaries) %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~colorQuantile("YlOrRd", shape_area)(shape_area),
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE)) %>% 
    addProviderTiles(providers$CartoDB.Positron)
  })

renderLeaflet({
      leaflet(divvy) %>% 
      addCircles(divvy$geometry) %>% 
      addTiles() %>%
      addCircleMarkers(data = divvy, divvy$geometry, 
                       radius = 3, popup = ~as.character(cntnt), 
                       color = ~pal(divvy$Status),
                       stroke = FALSE, fillOpacity = 0.8)%>%
      addLegend(pal=pal, values=bb_data$Category,opacity=1, na.label = "Not Available")%>%
      addEasyButton(easyButton(
        icon="fa-crosshairs", title="ME",
        onClick=JS("function(btn, map){ map.locate({setView: true}); }")))
        })
renderLeaflet({
      leaflet(cpsPR2018_sf) %>% 
      addCircles(cpsPR2018_sf$School_Type) %>% 
      addTiles() %>%
      addCircleMarkers(data = cpsPR2018_sf, cpsPR2018_sf$geometry, 
                       radius = 3, popup = ~as.character(cntnt), 
                       color = ~pal(cpsPR2018_sf$School_Type),
                       stroke = FALSE, fillOpacity = 0.8)%>%
      addLegend(pal=pal, values=bb_data$Category,opacity=1, na.label = "Not Available")%>%
      addEasyButton(easyButton(
        icon="fa-crosshairs", title="ME",
        onClick=JS("function(btn, map){ map.locate({setView: true}); }")))
        })
```

The Tables
=======================================================================
Select {.sidebar}
-----------------------------------------------------------------------
```{r}
#selectInput("data", "Choose a dataset:", choices = c("Divvy", "CTA Ridership", "CTA Stations", "Chicago SVI", "CPS Tiers", "Chicago Public Schoos", "Tax Increment Financing"))


conditionalPanel(
        'input.dataset === "Divvy Stations"',
        helpText("Click the column header to sort a column.")
      )
      conditionalPanel(
        'input.dataset === "CTA Ridership"',
        helpText("Click the column header to sort a column.")
      )
      conditionalPanel(
        'input.dataset === "CTA Stations"',
        helpText("Click the column header to sort a column.")
      )
      conditionalPanel(
        'input.dataset === "Chicago SVI"',
        helpText("Click the column header to sort a column.")
      )
      conditionalPanel(
        'input.dataset === "CPS Tiers"',
        helpText("Click the column header to sort a column.")
      )
      conditionalPanel(
        'input.dataset === "Chicago Public Schools"',
        checkboxGroupInput("show_vars", "Columns in dataset to show:",
                           names(cpsPR2018_sf), selected = names(cpsPR2018_sf))
      )
      conditionalPanel(
        'input.dataset === "Tax Increment Financing"',
        helpText("Click the column header to sort a column.")
      )
      conditionalPanel(
        'input.dataset === "Affordable Rental Housing"',
        helpText("Click the column header to sort a column.")
      )
      
downloadButton(outputId = "down", label = "Download the Data")

downloadHandler(filename = function() {
     paste('resultset-', Sys.Date(), '.csv', sep='')
   },
     content = function(file) {
     write.csv(subset_dataset(), file, row.names = FALSE)
   }
)
```

Output
-----------------------------------------------------------------------
### Data
```{r}
tabsetPanel(
        id = 'dataset',
        tabPanel("Divvy Stations", DT::dataTableOutput("mytable1")),
        tabPanel("CTA Ridership", DT::dataTableOutput("mytable2")),
        tabPanel("CTA Stations", DT::dataTableOutput("mytable3")),
        tabPanel("Chicago SVI", DT::dataTableOutput("mytable4")),
        tabPanel("CPS Tiers", DT::dataTableOutput("mytable5")),
        tabPanel("Chicago Public Schools", DT::dataTableOutput("mytable6")),
        tabPanel("Tax Increment Financing", DT::dataTableOutput("mytable7")),
        tabPanel("Affordable Rental Housing", DT::dataTableOutput("mytable8"))
      )

output$mytable1 <- DT::renderDataTable({
    DT::datatable(divvy, options(DT.options = list(scrollY="100vh")))
  })
  
output$mytable2 <- DT::renderDataTable({
    DT::datatable(cta_ridership, options = list(c(scrollY="200px", scrollX="300px", pageLength = 100)))
  })

output$mytable3 <- DT::renderDataTable({
    DT::datatable(cta_stations, options = list(c(scrollY="200px", scrollX="300px", pageLength = 100)))
  })

output$mytable4 <- DT::renderDataTable({
    DT::datatable(chicago_svi, options = list(c(scrollY="200px", scrollX="300px", pageLength = 100)))
  })

output$mytable5 <- DT::renderDataTable({
    DT::datatable(cps_tiers, options = list(c(scrollY="200px", scrollX="300px", pageLength = 100)))
  })

output$mytable6 <- DT::renderDataTable({
    DT::datatable(cpsPR2018_sf, options = list(c(scrollY="200px", scrollX="300px", pageLength = 100)))
  })

output$mytable7 <- DT::renderDataTable({
    DT::datatable(tif_sf, options = list(c(scrollY="200px", scrollX="300px", pageLength = 100)))
  })

output$mytable8 <- DT::renderDataTable({
    DT::datatable(tif_sf, options = list(c(scrollY="200px", scrollX="300px", pageLength = 100)))
  })
```

Codebook
=======================================================================
### Information Regarding the Data
<p><span>Datasets to be included in the package include (All accessed through SODA API):</span></p>
<ul>
<ul>
<li><span>Census Tracts 2010 (soon to be updated to 2020): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Census-Tracts-2010/5jrd-6zik"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Community Areas (current): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas-current-/cauq-8yn6"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Police Beats (current): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Beats-current-/aerh-rz74"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Police Districts (current): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Districts-current-/fthy-xz3r"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Wards (2015-): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Wards-2015-/sp34-6z76"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Zoning Districts (current): Accessed <a href="https://data.cityofchicago.org/Community-Economic-Development/Boundaries-Zoning-Districts-current-/7cve-jgbp"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Affordable Rental Housing Developments (points) (2019): Accessed </span><span>here</span><span> as a </span><span>json</span><span> file;</span></li>
<li><span>Chicago Public Schools - School Progress Reports SY1617 (points, with information regarding community areas): Accessed <a href="https://data.cityofchicago.org/resource/cp7s-7gxg.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>CTA - Ridership - Monthly Day-Type Averages &amp; Totals: Accessed <a href="https://data.cityofchicago.org/Transportation/CTA-Ridership-Bus-Routes-Monthly-Day-Type-Averages/bynn-gwxy"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>CTA - 'L' (Rail) Stations: Accessed <a href="https://data.cityofchicago.org/Transportation/CTA-L-Rail-Stations-Shapefile/vmyy-m9qj"><span>here</span></a> as a </span><span>zip</span><span> file (points, with community area information) (will have to figure out a way to unzip and manage this dataset, possibly a new R script);</span></li>
<li><span>Census Data - Selected socioeconomic indicators in Chicago, 2008 &ndash; 2012 (by community area): Accessed <a href="https://data.cityofchicago.org/resource/kn9c-c2s2.json"><span>here</span></a> as a </span><span>json</span><span> file. This dataset is outdated, but is the only one that shows community-area level data. It is possible to use CPS data directly from census bureau or IPUMS, but they are not as detailed to the community area level;</span></li>
<li><span>Divvy Bicycle Stations (points, with community area info): Accessed <a href="https://data.cityofchicago.org/resource/bbyy-e7gq.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>Tax Increment Financing (TIF) Projects (points): Accessed <a href="https://data.cityofchicago.org/Community-Economic-Development/Tax-Increment-Financing-TIF-Projects-Map/v3a3-hhqn"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>311 Service Requests - Vacant and Abandoned Buildings (points, with community area information): Accessed <a href="https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Vacant-and-Abandoned-Building/d9re-tmpw"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
</ul>
</ul>
<p><span>The above datasets are all original sets included in Clyde&rsquo;s Package, and I linked online API locations, and will update in the code. I also included an additional dataset below which I think is quite important:</span></p>
<ul>
<ul>
<li><span>Crimes - 2001 to Present (points, with temporal information, can aggregate by community area, police wards, beats): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Crimes-Map/dfnk-7re6"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>Divvy Trips (points, with beginning and end point info): Accessed <a href="https://data.cityofchicago.org/resource/fg6s-gzvg.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
</ul>
</ul>
<p><span>More datasets will be added after the completion of the first stage of project.</span></p>
<p><span>&nbsp;</span></p>


About
=======================================================================
My name is Ryan Wang, and I am a Master's student at the Univeristy of Chicago, concentrating on quantitative methods of social analyses. I created this flexdashboard app with RShiny code incorporated. This app uses the data and functionalities in the Chicago Package and it intends to be a tool for initial assessment of the Chicago data, aiming to quickly guide interests for further data discovery.

The "Chicago Package" is a continuation of Clyde Schwarb's Chicago Data Package (https://github.com/cschwab1/chicagodatapackage) project for GIS 3. Clyde is an undergraduate at the University of Chicago studying geography and environmental and urban studies. A year ago, Clyde initated this project to "provide a starting point for those interested in learning geocomputation with R using real Chicago data", and "building a larger collection of civic data for academic and amateur, journalist and researcher, non-profit and concerned citizens alike." Some most commonly used datasets were provided in the original package, along with initial processing code and vignettes from Clyde. I completed documentation of datasets and added two new datasets, Crime - from 2001 to present, and Divvy Trips.

I have further improved upon the package by providing complete data documentation and vignettes. I also added two new datasets and updated exisiting datasets to current June 2020 data.

The app is made into two sections, a spatial data explorer section ("The Maps") and a non-spatial data explorer section("The Tables"). Both uses a sidebar to select specific variables of interest, and has the functionality for downloading the data.

Copyright (c) 2019 Clyde
Copyright (c) 2020 Ryan
