---
title: "Chicago Data Explorer"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    social: menu
    author: 'Ryan Wang'
    theme: bootstrap
    # logo: www/73F1665C-0FE6-4183-8AD1-E91DB8EFAFA4.png
runtime: shiny
---
<style>
:root{
  font-family: 'Open Sans';
}
.navbar {
  background-color:#c7e5f3;
  border-color:#c7e5f3;
}

.section.sidebar {
  top: 61px;
}

.navbar-brand {
  color:black!important;
  font-weight: 500;
  
.sidebar {
  font-weight: 400;
}
</style>   
```{r setup, include=FALSE}
library(leaflet)
library(spData)
library(dplyr)
library(ChicagoPackage)
library(tmap)
library(sf)
```

The Maps
=======================================================================
Column {data-width=350}
-----------------------------------------------------------------------
### Select Dataset and Attributes
```{r}
selectizeInput("dataset", "Choose a dataset:", choices = c("Chicago Public Shools 2018" = cps_tiers, "Divvy" = divvy, "CTA Ridership" = cta_ridership, "CTA Stations" = cta_stations, "Crime" = crime_geo))
selectInput("variable", "Choose a variable:", choices = c("1", "2", "3"))
selectInput("by", "visualize by:", choices = c("ca_boundaries", "census_tracts_2010", "police_beats2019", "wards2019"))
selectInput("as", "show as:", choices = c("density dot", "dots", "Choropleth Map"))
sliderInput("year",
            "Year",
            min = 2001,
            max = 2018,
            step = 1,
            sep = "",
            value = 1)


```
```{r, eval=FALSE}
selectInput("dataset", "Choose a dataset:", choices = c("Chicago Public Shools 2018" = cpsPR2018_sf, "Divvy" = divvy, "CTA Ridership" = cta_ridership, "CTA Stations" = cta_stations))
selectInput("variable", "Choose a variable:", choices = c("1", "2", "3"))
selectInput("by", "visualize by:", choices = c(ca_boundaries, census_tracts_2010, police_beats2019, wards2019))
selectInput("as", "show as:", choices = c("density dot", "dots", "Choropleth Map"))
sliderInput("year",
            "Year",
            min = 2001,
            max = 2018,
            step = 1,
            sep = "",
            value = 1)

downloadButton(outputId = "down", label = "Download the Graph")
downloadHandler(filename = function() {
     paste('resultset-', Sys.Date(), '.csv', sep='')
   },
     content = function(file) {
     write.csv(subset_dataset(), file, row.names = FALSE)
   }
)
```

### Download the Plot
```{r}
radioButtons("picformat", "Select the Image Format to Download", list("png", "jpeg"), "")

downloadButton(outputId = "down", label = "Download the Graph")
downloadHandler(filename = function() {
     paste('resultset-', Sys.Date(), '.csv', sep='')
   },
     content = function(file) {
     write.csv(subset_dataset(), file, row.names = FALSE)
   }
)
```

Output {data-width=650}
-----------------------------------------------------------------------
### Output Map
```{r}

renderLeaflet({
  leaflet(ca_boundaries) %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~colorQuantile("YlOrRd", shape_area)(shape_area),
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE)) %>% 
    addProviderTiles(providers$CartoDB.Positron)
  })

```

The Tables
=======================================================================
Select {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("dataset", "Choose a dataset:", choices = c("Crime" = crime, "Divvy", "CTA Ridership", "CTA Stations"))
selectInput("variable", "Choose a variable:", choices = c("1", "2", "3"))
sliderInput("year",
            "Year",
            min = 2001,
            max = 2018,
            step = 1,
            sep = "",
            value = 1)
downloadButton(outputId = "down", label = "Download the Data")

downloadHandler(filename = function() {
     paste('resultset-', Sys.Date(), '.csv', sep='')
   },
     content = function(file) {
     write.csv(subset_dataset(), file, row.names = FALSE)
   }
)
```

Output
-----------------------------------------------------------------------
### Data
```{r}

renderTable({
  p<-read.csv("/Users/ryan/Desktop/chicagodatapackage-master/data-raw/AffordableHousingDevelopments.csv")
})
```

Codebook
=======================================================================
### Information Regarding the Data
<p><span>Datasets to be included in the package include (All accessed through SODA API):</span></p>
<ul>
<ul>
<li><span>Census Tracts 2010 (soon to be updated to 2020): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Census-Tracts-2010/5jrd-6zik"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Community Areas (current): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas-current-/cauq-8yn6"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Police Beats (current): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Beats-current-/aerh-rz74"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Police Districts (current): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Districts-current-/fthy-xz3r"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Wards (2015-): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Wards-2015-/sp34-6z76"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Boundaries - Zoning Districts (current): Accessed <a href="https://data.cityofchicago.org/Community-Economic-Development/Boundaries-Zoning-Districts-current-/7cve-jgbp"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>Affordable Rental Housing Developments (points) (2019): Accessed </span><span>here</span><span> as a </span><span>json</span><span> file;</span></li>
<li><span>Chicago Public Schools - School Progress Reports SY1617 (points, with information regarding community areas): Accessed <a href="https://data.cityofchicago.org/resource/cp7s-7gxg.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>CTA - Ridership - Monthly Day-Type Averages &amp; Totals: Accessed <a href="https://data.cityofchicago.org/Transportation/CTA-Ridership-Bus-Routes-Monthly-Day-Type-Averages/bynn-gwxy"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>CTA - 'L' (Rail) Stations: Accessed <a href="https://data.cityofchicago.org/Transportation/CTA-L-Rail-Stations-Shapefile/vmyy-m9qj"><span>here</span></a> as a </span><span>zip</span><span> file (points, with community area information) (will have to figure out a way to unzip and manage this dataset, possibly a new R script);</span></li>
<li><span>Census Data - Selected socioeconomic indicators in Chicago, 2008 &ndash; 2012 (by community area): Accessed <a href="https://data.cityofchicago.org/resource/kn9c-c2s2.json"><span>here</span></a> as a </span><span>json</span><span> file. This dataset is outdated, but is the only one that shows community-area level data. It is possible to use CPS data directly from census bureau or IPUMS, but they are not as detailed to the community area level;</span></li>
<li><span>Divvy Bicycle Stations (points, with community area info): Accessed <a href="https://data.cityofchicago.org/resource/bbyy-e7gq.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>Tax Increment Financing (TIF) Projects (points): Accessed <a href="https://data.cityofchicago.org/Community-Economic-Development/Tax-Increment-Financing-TIF-Projects-Map/v3a3-hhqn"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
<li><span>311 Service Requests - Vacant and Abandoned Buildings (points, with community area information): Accessed <a href="https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Vacant-and-Abandoned-Building/d9re-tmpw"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
</ul>
</ul>
<p><span>The above datasets are all original sets included in Clyde&rsquo;s Package, and I linked online API locations, and will update in the code. I also included an additional dataset below which I think is quite important:</span></p>
<ul>
<ul>
<li><span>Crimes - 2001 to Present (points, with temporal information, can aggregate by community area, police wards, beats): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Crimes-Map/dfnk-7re6"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
<li><span>Divvy Trips (points, with beginning and end point info): Accessed <a href="https://data.cityofchicago.org/resource/fg6s-gzvg.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
</ul>
</ul>
<p><span>More datasets will be added after the completion of the first stage of project.</span></p>
<p><span>&nbsp;</span></p>


About
=======================================================================
Building on variable health dashboard costs $10K, and it is absolutely crazy that something I can build in a day costs this much in real life!
