---
title: "README"
author: Ryan Wang
date: June 14, 2020
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chicago Data Package

 <!-- badges: start -->
 <!-- badges: end -->

My name is Ryan Wang, and I am a Master's student at the Univeristy of Chicago, concentrating on quantitative methods of social analyses. For this project, I further developed an open Chicago data package from a past GIS 3 class.

This is a continuation of Clyde Schwarb's Chicago Data Package (https://github.com/cschwab1/chicagodatapackage) project for GIS 3. Clyde is an undergraduate at the University of Chicago studying geography and environmental and urban studies. A year ago, Clyde initated this project to "provide a starting point for those interested in learning geocomputation with R using real Chicago data", and "building a larger collection of civic data for academic and amateur, journalist and researcher, non-profit and concerned citizens alike." Some most commonly used datasets were provided in the original package, along with initial processing code and vignettes from Clyde. I completed documentation of datasets and added two new datasets, Crime - from 2001 to present, and Divvy Trips.

Cleaned datasets are available in /data, and original data is included the unprocessed data in /data-raw. The data was primarily collected from the Chicago data portal (https://data.cityofchicago.org/), but also contains datasets from the CDC and Cook County. Additional documentation is available in /R. 

## Installation
The project is not yet available from CRAN, so you can download it directly from github:
```{r, eval=FALSE}
library(devtools)
install_github("ryanwyg/Chicago-Data-Package")
```
The installation might take several minutes depending on your internet connection, as some of the datasets in this pacakge are large.

## Datasets Included
 <p><span>Datasets to be included in the package include (All accessed through SODA API):</span></p>
 <ul>
 <ul>
 <li><span>Census Tracts 2010 (soon to be updated to 2020): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Census-Tracts-2010/5jrd-6zik"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 <li><span>Boundaries - Community Areas (current): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas-current-/cauq-8yn6"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 <li><span>Boundaries - Police Beats (current): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Beats-current-/aerh-rz74"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 <li><span>Boundaries - Police Districts (current): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Boundaries-Police-Districts-current-/fthy-xz3r"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 <li><span>Boundaries - Wards (2015-): Accessed <a href="https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Wards-2015-/sp34-6z76"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 <li><span>Boundaries - Zoning Districts (current): Accessed <a href="https://data.cityofchicago.org/Community-Economic-Development/Boundaries-Zoning-Districts-current-/7cve-jgbp"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 <li><span>Affordable Rental Housing Developments (points) (2019): Accessed </span><span>here</span><span> as a </span><span>json</span><span> file;</span></li>
 <li><span>Chicago Public Schools - School Progress Reports SY1617 (points, with information regarding community areas): Accessed <a href="https://data.cityofchicago.org/resource/cp7s-7gxg.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
 <li><span>CTA - Ridership - Monthly Day-Type Averages &amp; Totals: Accessed <a href="https://data.cityofchicago.org/Transportation/CTA-Ridership-Bus-Routes-Monthly-Day-Type-Averages/bynn-gwxy"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
 <li><span>CTA - 'L' (Rail) Stations: Accessed <a href="https://data.cityofchicago.org/Transportation/CTA-L-Rail-Stations-Shapefile/vmyy-m9qj"><span>here</span></a> as a </span><span>zip</span><span> file (points, with community area information) (will have to figure out a way to unzip and manage this dataset, possibly a new R script);</span></li>
 <li><span>Census Data - Selected socioeconomic indicators in Chicago, 2008 &ndash; 2012 (by community area): Accessed <a href="https://data.cityofchicago.org/resource/kn9c-c2s2.json"><span>here</span></a> as a </span><span>json</span><span> file. This dataset is outdated, but is the only one that shows community-area level data. It is possible to use CPS data directly from census bureau or IPUMS, but they are not as detailed to the community area level;</span></li>
 <li><span>Divvy Bicycle Stations (points, with community area info): Accessed <a href="https://data.cityofchicago.org/resource/bbyy-e7gq.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
 <li><span>Tax Increment Financing (TIF) Projects (points): Accessed <a href="https://data.cityofchicago.org/Community-Economic-Development/Tax-Increment-Financing-TIF-Projects-Map/v3a3-hhqn"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 <li><span>311 Service Requests - Vacant and Abandoned Buildings (points, with community area information): Accessed <a href="https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Vacant-and-Abandoned-Building/d9re-tmpw"><span>here</span></a> as a </span><span>geojson</span><span> file;</span></li>
 </ul>
 </ul>
 <p><span>The above datasets are all original sets included in Clyde&rsquo;s Package, and I linked online API locations, and will update in the code. I also included an additional dataset below which I think is quite important:</span></p>
 <ul>
 <ul>
 <li><span>Crimes - 2001 to Present (points, with temporal information, can aggregate by community area, police wards, beats): Accessed <a href="https://data.cityofchicago.org/Public-Safety/Crimes-Map/dfnk-7re6"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
 <li><span>Divvy Trips (points, with beginning and end point info): Accessed <a href="https://data.cityofchicago.org/resource/fg6s-gzvg.json"><span>here</span></a> as a </span><span>json</span><span> file;</span></li>
 </ul>
 </ul>
 <p><span>More datasets will be added after the completion of the first stage of project.</span></p>
 
#   

## The Chicago Data Explorer
This is a flexdashboard app with RShiny code incorporated. This app uses the data and functionalities in this package itself and it intends to be a tool for initial assessment of the Chicago data, aiming to quickly guide interests for further data discovery.

The app is made into two sections, a spatial data explorer section ("The Maps") and a non-spatial data explorer section("The Tables"). Both uses a sidebar to select specific variables of interest, and has the functionality for downloading the data.

## Known Issues

## Future Improvements
This project is ongoing and future functionalities will come soon, mainly being expansions in the data. You can come back and explore updates through the Data Explorer, which is constantly updated with new datasets from the city's data portal. I plan to expand further with more sections for different categories of data, just like how the data portal websites classifies the data ("Buildings", "Community", "Education" etc.) Please do check back for updates.
