---
title: "Exploring the Shapes"
author: "Ryan Wang"
date: "6-14-2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring the Shapes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
  # comment = "#>"
)
```

This is a tutorial using the ChicagoPackage. This tutorial mainly concerns the shapes in the package, namely the community areas, police wards and beats, and cencus tracts.
```{r}
### load the relevant packages
library(ChicagoPackage)
library(tmap)
library(sf)
library(rgeos)
library(rgdal)
library(spdep)
library(dplyr)
```

This Chicago Data Package contains different shapes, and this tutorial will show you what each individual shape looks like and how you might want to use them.

# Community Areas
```{r}
# Seeing the properties of the community area dataset
head(ca_boundaries)

# Plotting community areas, by the area
tm_shape(ca_boundaries) + tm_polygons("shape_area", palette = "Greens") +
  tm_layout(main.title = "Community Areas")
```
# Census Tracts
```{r}
# Again, properties
head(census_tracts_2010)

# Plotting census tracts, by their community areas
tm_shape(census_tracts_2010) + tm_polygons("commarea", palette = "Set3") +
  tm_layout(main.title = "2010 Census Tracts")
# We notice that there is a warning message, saying that there are more levels of community areas than the color palette, which is fine. As a result, some areas are grouped using the same color, resulting in larger areas.
```

# Police Beats
```{r}
# Same procedures as before
head(police_beats2019)
tm_shape(police_beats2019) + tm_polygons("beat", palette = "Accent") +
  tm_layout(main.title = "Police Beats")
```

# Police Wards
```{r}
# Same procedures as before
head(wards2019)
tm_shape(wards2019) + tm_polygons("ward", palette = "Paired") +
  tm_layout(main.title = "Police Wards")
```

